---
output: 
  pdf_document: 
    fig_crop: no
    fig_width: 7.5
    fig_height: 4.5
    latex_engine: xelatex
mainfont: Minion Pro
header-includes: 
  \usepackage{fancyhdr}
  \pagestyle{fancy}
  \fancyhf{ }
  \chead{DRAFT - 02/2020}
  \rfoot{\includegraphics[height=0.15in]{../_utils/DataHaven_Black.png}}
  \pagenumbering{gobble}
geometry: 
- top=0.75in
- right=0.5in
- bottom=0.75in
- left=0.5in
linespread: 1.1
subparagraph: yes
urlcolor: "blue"
params:
  geoid: "S004"
---

---
title: Connecticut State Legislative Profile, 2020
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.showtext = TRUE,
	message = FALSE,
	warning = FALSE,
	dpi = 300
)
```

```{r}
library(waffle)
library(patchwork)
```

```{r}
source("../_utils/utils.R")
theme_set(theme_din2())

dist_id <- params$geoid
chambers <- list(H = "house", S = "senate")
chamber <- chambers[[substr(dist_id, 1, 1)]]
is_senate <- chamber == "senate"
dist_num <- parse_number(dist_id)
dist_name <- sprintf("State %s District %s", str_to_title(chamber), dist_num)
dist_shortname <- paste0(toupper(substr(dist_id, 1, 1)), dist_num)
dist_midname <- paste("Dist.", dist_num)
dist_sf <- dists_sf[[chamber]] %>% 
  filter(id == substring(dist_id, 2)) %>%
  mutate(lbl = dist_midname)

reps <- jsonlite::read_json("../to_viz/legislators_meta.json", simplifyVector = TRUE)[[chamber]] 
dist_rep <- reps %>%
  mutate(rep = str_glue("{name} ({party})")) %>%
  filter(id == dist_id) %>%
  pull(rep)
towns <- jsonlite::read_json("../to_viz/town_lookup.json", simplifyVector = TRUE)[[chamber]][[dist_id]] %>%
  paste(collapse = ", ") %>%
  str_replace(", ([\\w\\s]+$)", " and \\1")
```



```{r}
prof <- readRDS("../output_data/acs_legislative_profile_2018.rds")
dist_prof <- prof %>%
  filter(level == "state" | name == dist_id) %>%
  mutate(name = fct_recode(name, !!!set_names(dist_id, dist_midname)))
totals <- dist_prof %>%
  filter(level == "district", group %in% c("total_pop", "total_households")) %>%
  distinct(group, estimate) %>%
  deframe() %>%
  as.list()
tbls <- split(dist_prof, dist_prof$topic, drop = TRUE)
cws <- readRDS("../output_data/cws_legislative_data_2018.rds") %>%
  mutate(level = ifelse(district == "Connecticut", "state", "district") %>% as_factor()) %>%
  mutate_at(vars(topic, question), as_factor) %>%
  mutate_if(is.factor, fct_drop) %>%
  filter(district %in% c("Connecticut", dist_id)) %>%
  mutate(topic = fct_relabel(topic, clean_titles),
         question = question %>% 
           fct_relabel(clean_titles) %>%
           fct_recode("Very good self-rated health" = "Very good self rate health", "Dental visit in past year" = "Dental visit past yr", "Local parks in good condition" = "Good cond of parks", "Feel local govt is responsive" = "Local govt responsive", "Think residents are able to obtain jobs" = "Suitable employment", "Good place to raise kids" = "Good to raise kids", "Adults with diabetes" = "Diabetes"),
         district = fct_recode(district, !!!set_names(dist_id, dist_midname)))
```

See data and interactive maps for all state Senate and House districts on DataHaven's Dashboard page (https://www.ctdatahaven.org/data-dashboard).

# `r dist_name`

* Represented by `r dist_rep`
* Covers all or parts of `r towns`
* Total population: `r comma(totals[["total_pop"]])` (`r comma(totals[["total_households"]])` households)



```{r distmaps, fig.width=7.5, fig.height=4.5}
# ct_bbox <- state_sf %>%
#   st_buffer(5e-2) %>%
#   st_bbox() %>%
#   setNames(c("left", "bottom", "right", "top"))
# ct_basemap <- get_stamenmap(ct_bbox, zoom = 10, maptype = "toner-lite")

state_map <- ggmap(ct_basemap) +
  geom_sf(data = dist_sf, color = bupu7[6], fill = bupu7[6], alpha = 0.5, size = 0.4, inherit.aes = FALSE) +
  geom_sf(data = state_sf, color = bupu7[7], fill = "transparent", size = 0.35, inherit.aes = FALSE) +
  coord_sf(ndiscr = 0) +
  labs(x = NULL, y = NULL)

dist_bbox <- dist_sf %>%
  st_buffer(1e-2) %>%
  st_bbox() %>%
  setNames(c("left", "bottom", "right", "top"))
dist_basemap <- get_stamenmap(dist_bbox, zoom = 13, maptype = "toner-lite")

dist_map <- ggmap(dist_basemap) +
  geom_sf(aes(fill = lbl, color = lbl, alpha = lbl), data = dist_sf, size = 0.8, inherit.aes = FALSE) +
  geom_sf(aes(fill = "Town boundaries", color = "Town boundaries", alpha = "Town boundaries"), data = cwi::town_sf, size = 0.5, inherit.aes = FALSE) +
  coord_sf(ndiscr = 0) +
  labs(x = NULL, y = NULL, fill = NULL, color = NULL, alpha = NULL) +
  scale_fill_manual(values = c(bupu7[5], "transparent")) +
  scale_color_manual(values = c(bupu7[5], bupu7[7])) +
  scale_alpha_manual(values = c(0.5, 0.00)) +
  theme(plot.margin = margin(6, 6, 6, 12, "pt"),
        legend.position = "bottom",
        legend.key.height = unit(0.8, "lines"),
        legend.key.width = unit(1.1, "lines"),
        legend.text = element_text(margin = margin(0, 1, 0, 0, "lines")))

# cowplot::plot_grid(state_map, dist_map,
#                    nrow = 1)
state_map +
  dist_map +
  plot_layout(nrow = 1)
```



## Demographics

```{r age, fig.height=3, fig.width=7.5}
# for top corner, get max x, then max y
pop_age <- tbls$age %>%
  select(-share) %>%
  pivot_wider(names_from = group, values_from = estimate) %>%
  mutate(ages18_64 = ages18plus - ages65plus) %>%
  select(-ages18plus) %>%
  pivot_longer(-c(topic:name), names_to = "group") %>%
  mutate(group = fct_relevel(group, "ages65plus", after = Inf)) %>%
  arrange(group) %>%
  group_by_at(vars(-group, -value)) %>%
  calc_shares(value = value) %>%
  ungroup() %>%
  mutate(group = fct_relabel(group, clean_ages)) %>%
  filter(group != "Total pop") %>%
  mutate_if(is.factor, fct_drop)

age_pal <- setNames(pastel[c(2, 3, 8)], levels(pop_age$group))

p_age <- pop_age %>%
  group_by(name) %>%
  mutate(share = round_sum100(share * 100)) %>%
  ggplot() +
  geom_waffle(aes(fill = group, values = share), flip = TRUE, radius = unit(2, "pt"), size = 0.6, color = "white") +
  coord_fixed() +
  facet_wrap(vars(name))  +
  scale_fill_manual(values = age_pal)

age_lbls <- ggplot_build(p_age)$data[[1]] %>%
  inner_join(enframe(age_pal, name = "group", value = "fill"), by = "fill", suffix = c(".y", "")) %>%
  group_by(PANEL, group) %>%
  mutate(sqrs = n()) %>%
  filter(x == max(x)) %>%
  filter(y == max(y)) %>%
  distinct(PANEL, group, .keep_all = TRUE) %>%
  ungroup() %>%
  select(panel = PANEL, group, x, y, sqrs) %>%
  mutate(panel = as.numeric(as.character(panel))) %>%
  # bind_cols(pop_age) %>%
  left_join(pop_age %>% mutate(panel = as.numeric(name)), by = c("group", "panel")) %>%
  mutate(lbl = str_glue("{group}: {percent(share, accuracy = 1)}"),
         offset = ifelse(sqrs == 1, 0.6, 0))


p_age +
  geom_text(aes(x = x, y = y + offset, label = lbl),
            data = age_lbls,
            hjust = 1, size = 2.8, fontface = "plain", nudge_x = 0.25) +
  # scale_x_continuous(expand = expansion(add = 0.1)) +
  scale_y_continuous(expand = expansion(add = c(0.1, 0.25))) +
  theme_din2(ygrid = FALSE) +
  theme(axis.text = element_blank(),
        legend.position = "none",
        strip.text = element_text(face = "plain", vjust = 0, hjust = 0.05)) +
  labs(x = NULL, y = NULL, title = "Population by age, 2018")
```

```{r race, fig.height=3, fig.width=7.5}
pop_race <- tbls$race %>%
  select(-share) %>%
  group_by_at(vars(topic:name)) %>%
  calc_shares() %>%
  ungroup() %>%
  mutate(group = group %>% 
           fct_relevel("total_pop", "white") %>%
           fct_relabel(clean_titles)) %>%
  arrange(name, group) %>%
  filter(group != "Total pop") %>%
  mutate_if(is.factor, fct_drop)

race_pal <- setNames(pastel[c(1, 3, 2, 4)], levels(pop_race$group))

p_race <- pop_race %>%
  group_by(name) %>%
  mutate(share = round_sum100(share * 100)) %>%
  ggplot(aes(fill = group, values = share)) +
  geom_waffle(flip = TRUE, radius = unit(2, "pt"), color = "white", size = 0.6) +
  coord_fixed() +
  scale_fill_manual(values = race_pal) +
  facet_wrap(vars(name)) 

race_lbls <- ggplot_build(p_race)$data[[1]] %>%
  inner_join(enframe(race_pal, name = "group", value = "fill"), by = "fill", suffix = c(".y", "")) %>%
  group_by(PANEL, group) %>%
  mutate(sqrs = n()) %>%
  filter(x == max(x)) %>%
  filter(y == max(y)) %>%
  distinct(PANEL, group, .keep_all = TRUE) %>%
  ungroup() %>%
  select(panel = PANEL, group, x, y, sqrs) %>%
  mutate(panel = as.numeric(as.character(panel))) %>%
  # bind_cols(pop_age) %>%
  left_join(pop_race %>% mutate(panel = as.numeric(name)), by = c("group", "panel")) %>%
  mutate(lbl = str_glue("{group}: {percent(share, accuracy = 1)}"),
         offset = ifelse(sqrs == 1, 0.6, 0))

p_race +
  geom_text(aes(x = x, y = y + offset, label = lbl),
            data = race_lbls,
            hjust = 1, size = 2.8, fontface = "plain", nudge_x = 0.25) +
  # scale_x_continuous(expand = expansion(add = c(0, 3))) +
  scale_y_continuous(expand = expansion(add = c(0.1, 0.25))) +
  theme_din2(ygrid = FALSE) +
  theme(axis.text = element_blank(),
        legend.position = "none",
        strip.text = element_text(face = "plain", vjust = 0, hjust = 0.05)) +
  labs(x = NULL, y = NULL, title = "Population by race, 2018")
```



## Households and income

```{r households, fig.width=7.5, fig.height=2.5}
housing_income <- tbls[c("housing", "income")] %>%
  bind_rows() %>%
  mutate(group = as_factor(group)) %>%
  filter(!is.na(share)) %>%
  mutate(group = group %>%
           fct_relabel(clean_titles) %>%
           fct_recode("Homeownership" = "Owner occupied", "Low-income" = "Low income") %>%
           fct_relabel(paste, "rate"))

housing_income %>%
  mutate(per_lbl = percent(share, accuracy = 1),
         # num_lbl = str_glue("\u00a0\n({comma(estimate)})")
         lbl = ifelse(level == "district", str_glue("{per_lbl}\n({comma(estimate)})"), per_lbl)
         ) %>%
  mutate(ratio = share / max(share),
         small = ratio < 0.13,
         offset = ifelse(small, 1, -1) * 1e-2) %>%
  # mutate(lbl = paste0(bold(per_lbl), "\n(", num_lbl, ")")) %>%
  ggplot(aes(x = name, y = share, fill = level)) +
  facet_wrap(vars(group), nrow = 1) +
  geom_col(width = 0.8) +
  geom_text(aes(label = lbl, y = share + offset, color = small, vjust = as.numeric(!small)), lineheight = 0.95) +
  # geom_text(aes(label = num_lbl, y = share + offset, color = small, vjust = as.numeric(!small)),
  #           data = . %>% filter(level == "district"), 
  #           fontface = "plain", lineheight = 1.0) +
  scale_y_barcontinuous(breaks = NULL) +
  scale_x_discrete(expand = expansion(mult = 0.7)) +
  scale_fill_manual(values = c(state = "gray60", district = bupu7[5])) +
  scale_color_manual(values = c(`TRUE` = "gray40", `FALSE` = "white")) +
  theme(legend.position = "none",
        strip.text = element_text(face = "bold")) +
  labs(x = NULL, y = NULL)
```


```{r eval=is_senate, results='asis'}
pander::pandoc.header("Community well-being", 2)
```



```{r cws, fig.height=5, fig.width=7.2, eval=is_senate}
cws %>%
  arrange(topic) %>%
  mutate(question = fct_inorder(question)) %>%
  group_by(question) %>%
  mutate(offset = max(value) / 40) %>%
  ggplot(aes(x = district, y = value, fill = level)) +
  geom_col(width = 0.8) +
  geom_text(aes(label = percent(value, accuracy = 1), y = value - offset), color = "white", vjust = 1) +
  scale_y_barcontinuous(breaks = NULL) +
  scale_x_discrete(expand = expansion(mult = 0.6)) +
  scale_fill_manual(values = c(state = "gray60", district = bupu7[5])) +
  facet_wrap(vars(question), ncol = 5, scales = "free_y", labeller = labeller(.cols = label_wrap_gen())) +
  theme(legend.position = "none",
        strip.text = element_text(face = "bold", vjust = 1),
        panel.spacing.y = unit(0.25, "in")) +
  labs(x = NULL, y = NULL)
```

## Sources

```{r results="asis"}
if (is_senate) {
  cat("DataHaven analysis (2020) of:\n", "* US Census Bureau. (2018). American Community Survey 2018 5-year estimates.", "* DataHaven (2015 and 2018). Community Wellbeing Survey.", sep = "\n")
} else {
  cat("DataHaven analysis (2020) of US Census Bureau. (2018). American Community Survey 2018 5-year estimates.")
}
```


---

\linespread{1}
DataHaven

129 Church Street, Suite 605

New Haven, CT 06510

(203) 500-7059 / info@ctdatahaven.org

[ctdatahaven.org](http://www.ctdatahaven.org)

DataHaven is a non-profit organization with a 25-year history of public service to Greater New Haven and Connecticut. Our mission is to improve quality of life by collecting, sharing, and interpreting public data for effective decision making. DataHaven is a formal partner of the National Neighborhood Indicators Partnership of the Urban Institute in Washington, DC.